# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(linear_algebra).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


## Installation

`rTorch` is available in GitHub only at this moment.

Install `rTorch` with: 

`devtools::install_github("f0nzie/rTorch")`

Before start running `rTorch`, install a Python Anaconda environment first. 

1. Create a conda environment with `conda create -n myenv python=3.7`

2. Activate the new environment with `conda activate myenv`

3. Install PyTorch related packages with:  

`conda install python=3.6.6 pytorch-cpu torchvision-cpu matplotlib pandas -c pytorch`


Now, you can load `rTorch` in R or RStudio.

The automatic installation, like in `rtensorflow`, may be available later.

**Note.** `matplotlib` and `pandas` are not really necessary, but I was asked if `matplotlib` or `pandas` would in PyTorch, that I decided to put them for testing and experimentation. They both work.



## Matrices and Linear Algebra

There are five major type of Tensors in PyTorch

```{r}
library(rTorch)

bt <- torch$ByteTensor(3L, 3L)
ft <- torch$FloatTensor(3L, 3L)
dt <- torch$DoubleTensor(3L, 3L)
lt <- torch$LongTensor(3L, 3L)
Bt <- torch$BoolTensor(5L, 5L)

ft
dt
Bt
```

A 4D tensor like in MNIST hand-written digits recognition dataset:

```{r}
mnist_4d <- torch$FloatTensor(60000L, 3L, 28L, 28L)

# size
mnist_4d$size()

# length
length(mnist_4d)

# shape, like in numpy
mnist_4d$shape

# number of elements
mnist_4d$numel()
```

A 3D tensor:

```{r}
ft3d <- torch$FloatTensor(4L, 3L, 2L)
ft3d
```

```{r}
# get first element in a tensor
ft3d[1, 1, 1]
```


```{r}
bt
# [torch.ByteTensor of size 3x3]
```

```{r}
ft
# [torch.FloatTensor of size 3x3]
```


```{r}
# create a tensor with a value
torch$full(list(2L, 3L), 3.141592)
```


