# Bookdown tips {-}

## Label references {-}
You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(linearalgebra).

> Note. For the labels, do not use underscore `_` or dash `-`; they will confuse bookdown referencing. Example: avoid `linear_algebra` or `linear-algebra` as labels. Preferrably, use one word.

## Figures and Tables with references {-}
Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

## Long captions {-}
If you have a **long caption** with Latex included maybe is a good idea to write the caption as a separate paragraph, outside the `Rmarkdown` chunk. Just like this:


    (ref:ALongCaption) **A base plot of pressure vs. temperature**. This plot   
    shows the pressure readings at different temperatures. The initial temperature,   
    $t_0$, is nearer to the intersection of the axis.

Then, in the caption reference copy the tag `(ref:ALongCaption)` enclosing it in quotes or single quotes.

(ref:ALongCaption) **A base plot of pressure vs. temperature**. This plot shows the pressure readings at different temperatures. The initial temperature, $t_0$, is nearer to the intersection of the axis.

```{r fig-long-caption, fig.cap='(ref:ALongCaption)', out.width='80%', fig.asp=.75, fig.align='center', echo=FALSE}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

With tables, do the same with long captions.

    (ref:TabLongCaption) The following table shows the numerical data for the
    sepal and petal measurements.

And then insert the tag reference where the caption should go enclosed in quotes or single quotes.

(ref:TabLongCaption) The following table shows the numerical data for the sepal and petal measurements.

```{r tab-long-caption, tidy=FALSE, echo=FALSE}
knitr::kable(
  head(iris, 20), caption = '(ref:TabLongCaption)',
  booktabs = TRUE
)
```

## Citations and Bibiliography {-}
You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].